\begin{abstractpage}

\begin{abstract}{romanian}

  Computerele moderne folosesc tehnici de optimizare precum \emph{execuție
  out-of-order} și \emph{branch prediction}. \emph{Meltdown} și \emph{Spectre}
  sunt două atacuri care exploatează efectele secundare apărute la nivel
  microarhitectural în urma optimizărilor menționate. Prin intermediul acestora
  un atacator poate citi date private din zone arbitrare din memorie, fară
  privilegii și fară a exploata niciun bug de natură software. \emph{Intel},
  \emph{AMD} și \emph{ARM} au fost forțate în urma divulgării acestor atacuri
  să își schimbe designul procesoarelor în încercarea de a mitiga
  vulnerabilitățile la nivel hardware. În ciuda soluțiilor implementate, la
  jumătatea anului 2022, \emph{Spectre} afectează în continure majoritatea
  computerelor din lumea întreagă și rămâne un pericol pentru utilizatori și un
  subiect de mare interes pentru cercetători. În această lucrare vor fi
  prezentate particularitățile celor două atacuri. În final voi ilustra și
  rezultatele experimentelor personale sub forma unei implementari
  demonstrative, diferită de cea descrisă în materialul sursă
  \cite{spectre2019}, a unui atac de tip \emph{Spectre}, prin care un atacator
  poate citi întreg spatiul de memorie al unui alt proces.

\end{abstract}

\begin{abstract}{english}
  
  Modern computers are equiped with features such as \emph{out-of-order
  execution} and \emph{branch prediction}, which are used to reduce CPU idle
  time and improve performance. \emph{Meltdown} and \emph{Spectre} are two cyber
  attacks that exploit microarchitectural side-effects which appear as a result
  of such optimization techniques being used. An attacker can read private
  data of the vicim at arbitrary locations in memory, without exploiting any
  software bug. \emph{Intel}, \emph{AMD} and \emph{ARM} were forced to redesign
  their CPUs in order to mitigate the risks posed by \emph{Meltdown} and
  \emph{Spectre}. Despite deployed mitigations, in the second half of 2022,
  most computers in the world are vulnerable to variations of \emph{Spectre}
  attacks, billions of users being at risk. This class of attacks remains a
  subject of great interest for researchers in the field of security. In this
  work, the technicalities and implications of both attacks will be covered. In
  the end, the results of my own experiements will be presented, as a proof of
  concept for a \emph{Spectre} attack. This implementation is different from
  the one in the research paper \cite{spectre2019} and allows an attacker to
  read the whole address space of another process.

\end{abstract}

\end{abstractpage}
